# Lecture 10 Prescriptive Analytics I: Linear Optimization
# Linear Optimization Solver in R

#### Preparation
load necessary library and packages for the analysis. install any if necessary
'car' stands for Companion for Applied Regression

```{r load-libraries, echo=TRUE,dependencies = TRUE, warning = FALSE, message = FALSE}
library(lpSolve)
library(animation)
library(ggplot2)
library(wooldridge)
```

# lpSolve for Linear Optimization in R

### Example A: Farmer Jean's Problem

Problem Description:  
Maximize total profit using decision variables X1, X2:     Profit = 0.15 x1 + 0.40 x2  
Subject to:  
Budgetary constraints                                      0.20 x1 + 0.70 x2 ≤ 100  
Land Constraints                                           1 x1 + 1 x2 ≤ 200  
Non-negativity Constraints                                 x1 ≥ 0; x2 ≥ 0  

```{r example A, echo=TRUE}
# defining parameters 
objective_function = c(0.15, 0.40)
constraint_mat = matrix(c(0.20, 0.70, 1, 1), ncol = 2, byrow = TRUE)
constraint_dir = c('<=', '<=')
constraint_rhs = c(100, 200)
# solving the linear problem
lp_solution = lp(direction = 'max', objective_function, 
              constraint_mat, constraint_dir, constraint_rhs,
              compute.sens = TRUE)
# display the solution of linear problem: solution stored in 'lp_obj$solution'
print(lp_solution$solution)
# display the value of objective function at optimal solution
print(lp_solution)

# display range of objective coefficients where current solution is valid.
print(lp_solution$sens.coef.from); print(lp_solution$sens.coef.to);
range_objcoef = cbind(lp_solution$sens.coef.from, lp_solution$sens.coef.to)
rownames(range_objcoef) = c('coef x1', 'coef x2')
colnames(range_objcoef) = c('from', 'to')
print(range_objcoef)
# display the shadow prices of constraints in sensitivity analysis
print(lp_solution$duals)
```

### Example B: Vegan's Utility Maximization Problem

Problem Description:  
Maximize utlity choosing x        utility = 250 x1 + 225 x2 + 300 x3  
Subject to:  
Budget constraints                7 x1 + 5 x2 + 8 x3 ≤ 60  
Storage constraints               15 x1 + 30 x2 + 40 x3 ≤ 200  
Diet constraints                  x1 ≤ 7  
Non-negativity constraints        x1 ≥ 0; x2 ≥ 0; x3 ≥ 0  

```{r example B, echo=TRUE}
# first define all parameters
objective.fn <- c(250, 225, 300)
const.mat <- matrix(c(7, 5, 8, 15, 30, 40, 1, 0, 0),
                    ncol=3 , byrow=TRUE)
const.dir <- c("<=", "<=", "<=")
const.rhs <- c(60, 200, 7)
# then solve model
lp.solution <- lp("max", objective.fn, const.mat,
                  const.dir, const.rhs, compute.sens=TRUE)
# optimal solution for consumption profile
print(lp.solution$solution)
# objective fn value at optimal solution
print(lp.solution)

# display range of objective coefficients where current solution is valid.
print(lp.solution$sens.coef.from); print(lp.solution$sens.coef.to);
range.objcoef = cbind(lp.solution$sens.coef.from, lp.solution$sens.coef.to)
rownames(range.objcoef) = c('coef x1', 'coef x2', 'coef x3')
colnames(range.objcoef) = c('from', 'to')
print(round(range.objcoef, 1))

## display the shadow prices of constraints in sensitivity analysis
print(lp.solution$duals)
```