---
title: "Exam Testing"
output: html_document
---

```{r setup, include=FALSE}
# load required packages 
# library("xxxx")
library("Rmisc") # For computation of CI, summarySE (plotting of error bar)
library("tidyr") #spread, gather etc
library("rpivotTable")
library("knitr") #knitting
library("psych") #describe, corr.test()
library(readxl)
library(viridis)
library("rcompanion") #this package is required for transformTukey function
library("rstatix") # for Welch ANOVA & Games Howell test
library("dplyr") #need to call the libary before you use the package
library("TTR") #SMA
library(tsbox) #change df to ts model
library(ggplot2)
#car :: vif () #check for multicollinearity
#dplyr :: lag #for time series forcasting
```

## Some Formulas:
<p style="color:blue">
Mean use T-test, Proportion Z-test  
test statistic z for proportion: (proportion-probability) / sqrt(probability * (1-probability)/n)  
test statistic t for comparing means: (new mean-old mean)/(sd/sqrt(n))  
z critical value: qnorm(0.05) # 5% level of significance, one tailed  
t critical value: qt(0.975,df=n-1)  
confidence interval for mean: mean +- `test statistic t` * (sd/sqrt(n)) OR use CI()  
prediction interval: mean +- `test statistic t` * (sd * sqrt(1+(1/n))  
confidence interval for proportions: p +- `test statistic z`  * (sqrt(p*(1-p)/n)) 

Finding probability : pnorm()
</p>


```{r testing}
40.97045 - (qt(0.05, df = 199*20.32906*sqrt(1+1/200)))
40.97045 + (qt(0.05, df = 199*20.32906*sqrt(1+1/200)))




```


